---
title: 'Individual Assignment #3'
output:
  html_document:
    df_print: paged
---
***
<center> 
### Regression with ARIMA Errors Laboratory.
#### (Total 40 pts.)
#### Due: Dec. 2 (before midnight)
</center>
***

As a newly minted MSBA you start working for a large resort hotel managing 1,877 rooms.  A tour operator that you frequently do business with has requested a block of 60 rooms for Aug. 22.  As this operator will pay a discounted rate for each room, your manager is uncertain if she should give this block of rooms to the tour operator, and she  has asked you to provide model-based support for this decision.

Although the price is discounted and it may prevent you from selling some of these rooms at the regular price, if you reject this request, any empty rooms will not generate any revenue and you would have incurred an opportunity cost by not selling these rooms.

After judicious consideration of the economic trade offs in the News vendor model, you and your manager determine that the block should be assigned to the tour operator only if the probability of your organic demand (excluding the room block in question) is smaller than 67% (i.e., you are on track to achieve a 67% service level).

To address the booking question you examine the following data set and proceed to develop a demand forecast model.


The data set **Booking Exercise.csv** consists of daily observations of the following six variables:

* **DATE**: Calendar date corresponding to each observation.

* **DOW**: Index (1-7) to indicate the day of the week each **DATE** corresponds to. This is redundant and it is eliminated below.

* **DEMAND**: Number of rooms actually sold on each **DATE**.

* **TUESDAY.BOOK**: Number of rooms booked on the Tuesday of the previous week for each day of the forthcoming week.  This variable is used as an input to inform the forecasts of **DEMAND** for each day of the forthcoming week.

* **PICKUP.RATIO**: This is a calculated variable obtained as **PICKUP.RATIO = DEMAND/TUESDAY.BOOK** historically as **DEMAND** is observed. Because of this is a calculated relationship you can use either **PICKUP.RATIO** or **TUESDAY.BOOK** but you cannot use both variables simultaneously in a model to predict **DEMAND**. 

* **DOW.INDEX**: This is a pre-calculated indicator of day-of-the-week demand intensity.  This is obtained by the Hotel using information beyond what it is contained in this data set.

The code below reads the data set which includes up to the latest demand information (August 17) and prepares the data for the modeling process:

```{r}
library(fpp3)
library(tseries)
X <- read.csv("Booking Exercise.csv") %>%
  mutate(DATE = ymd(DATE)) %>%
  as_tsibble(index = DATE)


Fcst.X <- X %>% filter(DATE >= ymd("2010-08-18")) #forecastable data
X <- X %>% filter(DATE <= ymd("2010-08-17")) #historic data
```


1. (15 pt.) Fit (a) an ETS, (b) a plain ARIMA, (c) a linear regression model to predict **DEMAND** as a function of **TUESDAY.BOOK**, (c) a non-seasonal regression model with ARIMA errors using **TUESDAY.BOOK** and **DOW.INDEX** as explanatory variables, and (d) a seasonal regression model with ARIMA errors using only **TUESDAY.BOOK** as an explanatory variable.  You may use automatic model selection to fit an initial model, but especially in cases (c) and (d) you should modify the automatically fitted models when appropriate.
```{r}
#take a look at the data 
X %>% gg_tsdisplay(DEMAND, plot_type = "partial") +
  labs(title="X data", y="")
#differencing needed?
X %>% features(DEMAND, unitroot_nsdiffs) #seasonal differences needed = 0
X %>% features(DEMAND,unitroot_ndiffs) #differences needed = 0

m <- X %>%  
  model(auto.ETS = ETS(DEMAND), 
        auto.ARIMA = ARIMA(DEMAND), 
        lr = TSLM(DEMAND ~ TUESDAY.BOOK), #LR
        lr.ARIMA = ARIMA(DEMAND ~ TUESDAY.BOOK + DOW.INDEX), #non-seasonal LR with ARIMA errors
        lr.ARIMA.seasonal = ARIMA(DEMAND ~ TUESDAY.BOOK + PDQ(2,0,0))) #seasonal LR w/ ARIMA errors
```
2. (5 pt.) Which of the models are valid?
Based on the ljung_box test, although the linear regression model has the lowest information criteria values, it is not a valid model because it's p-value is smaller than .05. The residual plots are all a bit messy besides the non-seasonal LR with ARIMA errors model, the residuals of this model are normally distributed around zero. 
```{r}
#large p-value = it is a valid model. 
#If the p value is greater than 0.05 then the residuals are independent, which we want, and the model to be correct. If you simulate a white noise time series using the code below and use the same test for it then the p value will be greater than 0.05.
m %>% augment() %>%
  features(.resid, ljung_box, lag = 60)

m %>% select(auto.ETS) %>% gg_tsresiduals()
m %>% select(auto.ARIMA) %>% gg_tsresiduals()
m %>% select(lr) %>% gg_tsresiduals()
m %>% select(lr.ARIMA) %>% gg_tsresiduals() #best performing model
m %>% select(lr.ARIMA.seasonal) %>% gg_tsresiduals()
```
3. (5 pt.) Using the information criteria (AICc and BIC) what is the best model?

lr.ARIMA
Of the valid models, so not including linear regression, the non-seasonal linear regression with ARIMA errors model has the lowest information criteria values. 
```{r}
m %>% glance() %>% select(.model, AIC, AICc, BIC)
```
4. (5 pt.) Use the **Fcst.X** data set created above to produce a demand forecast for the forthcoming week (i.e., Aug, 18 - 24) showing the default 80% and 90% confidence intervals.
```{r}
fit <- m %>% select(lr.ARIMA)
forecast(fit, Fcst.X) 

forecast(fit, Fcst.X)  %>%
  autoplot(X) +
  labs(title="Forecasted Demand: Hotel Rooms")
```

5. (10 pt.) What is your recommendation for your manager regarding the tour operator block?  Should your manager sell the block of 60 rooms at a discounted price?  Justify your answer.
```{r}
1839/1877
```

Our August 22nd forecasted demand is 1839 rooms, which puts us at a service level of 97.9%. Since our forecasted demand shows we will be close to sold out on August 22nd, we should not give this client the block of 60 rooms. 

Something to consider is the accuracy of our model, the forecasted demand for August 21st is much lower than August 22nd, so if we are even a day off on our forecast then we have potential lost out on additional revenue. While we would rent the rooms at a discounted rate, sometimes confirmed sales is better than projected sales that do not end up occuring. 

With that being said, based on our best model chosen (lr.ARIMA) we would not book the block of 60 rooms because natural demand exceeds the set 67% service rate. 